---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

I chose to analyze the dataset that contained information about items checked out at least 10 times a month from 2017 - 2023. I thought this would be really interesting to look at to see what books have been popular within the Seattle population for a longer period of time. By understanding this information, I can better illustrate which types of media and of which subject category are most popular within people living in this area. This information can be utilized in the future to better inform what types of products the library invests in and how to improve the products and services based on what people typically enjoy.  

### Summary Information 

Overall, this dataset has 816,354  items checked out including ebooks, books, magazines, movies, and comics, to name a few. The most common method of checking out was by using OverDrive, a library company that digitally distributes ebooks, audiobooks, and online magazines. Over the course of 2017 - 2023, the item that was checked out the most number of times was "So You Want to Talk about Race" by Ijeoma Oluo. This is a non - fiction book published in 2018 that discusses race in contemporary America. Considering the vast amounts of protests related to police brutality and Black Lives Matter that began in 2020, I believe that this makes sense. However, in contrast to the genre of the most commonly checked out item, the category that had the most number of items checked out within it is comedy, spoofs, and satire. In a way, this also makes sense, as when the Covid - 19 pandemic began, people turned towards comedy and humor in order to cope with quarantine. In addition, the publisher that made the most number of checked out items was 'Entertainment One Films Canada', which was founded in 2007 and purchased by Hasbro in 2019. Finally, in 2017 the most common type of media checkout was videocasettes and video discs, while the most common type of media checkout in 2023 is ebooks. 

### The Dataset

This data is collected by the Seattle Public Library, the public library system serving Seattle, Washington that was founded in 1890. This dataset includes information about items checked out at least 10 times per month from 2017 - 2023. The checkout items include items like ebooks, books, magazines, movies, CDs, television dics, and audiobooks, though this is not an exhaustive list. This data was collected through the Open Data program generated by the City of Seattle. This program aims to provide open access to data with the hope of improving the quality of life for residents, promoting economic development and enrichment, and to identify opportunities to better serve the public. When looking at this data, I do not believe that there are any ethical questions that need to be addressed. However, I think it is important to think critically when it comes to implementing the conclusions that are generated using this data. When utilizing this data and the conclusions, it is important to think about how library resources are utilized by different groups of people, and implement policies and products that continue to be equitable to all Seattle residents. Finally, the possible limitations that come with this data include the inability to see at what point in the month people checked out items, whether that be at the beginning, middle, or end of the month. 

### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

I chose to look at trends of how the total number of items checked out from the Seattle Public Library has changed over time. I was interested in this because I was curious to see if the number of checkouts decreased when the pandemic began and people had less access to the physical products from the library. The graph shows that this is true, as there is a sharp decline in the number of checkouts from 2019 to 2020, and then a slight increase and subsequent decrease after. 

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, code = readLines("Chart1.R")}

```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

I was interested in looking at the total number of checkouts per item type, in order to determine what the most popular type of checkout item is. Not surprisingly, this graph shows that the most checked out item from the library is physical books, followed closely by audiobooks. Next, comic books and CR (what I believe to mean 'community room') are the next group of most checked out, followed by atlas or maps. 

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("Chart2.R")}

```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

For my last graph, I was curious to see what the top 10 most checked out books were from the library, because I really enjoy reading and was interested to see what types of books are most popular within the Seattle area. 

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("Chart3.R")}

```


